On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-03-27 manuscripts/PNAS/chazthm_pnas.tex
chaz

% Please include corresponding author, author contribution and author declaration information
\authordeclaration{The authors declare no conflict of interest.}
%\equalauthors{\textsuperscript{1}C.G. proved Theorem 1 and C.H. proved Theorem 2.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: cjg@berkeley.edu, chillar@msri.org}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Sparse linear coding, dictionary learning, compressed sensing}

\begin{abstract}
Dictionary learning methods for sparse linear coding have exposed underlying structure in many kinds of natural signals.
However, universal theorems guaranteeing the statistical consistency of inference in this model are lacking.
Here, we prove that for almost all diverse enough datasets generated from the sparse coding model, latent dictionaries and sparse codes are uniquely determined up to an error commensurate with measurement noise.
Applications are given to data analysis, neuroscience, and engineering.
\end{abstract}

@@

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{D}ictionary learning methods for sparse coding have been under active development ever since their application to natural images brought about an exciting discovery in computational neuroscience: that simple-cell response properties emerge from efficient coding principles \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent}. Much of the subsequent work in dictionary learning has come out of the signal processing and machine learning communities, which have found sparse representations to be a powerful tool for the manipulation of signals lacking much known analytic structure (see \cite{Zhang15} for a comprehensive review). This motivates the data-driven approach to inferring optimal dictionaries for a wide variety of signal classes for which, traditionally, representations have been constructed using methods of harmonic analysis (e.g., the Fourier basis, wavelets, curvelets, etc.). A popular formulation of the idea is to find a set of $m$ `elementary' waveforms, at most $k$ of which need be linearly combined to represent each $n$-dimensional signal in a dataset of size $N$, typically where $k < m \ll N$. 
%\dropcap{D}ictionary learning methods for sparse coding have been under active development ever since their application to natural images brought about an exciting discovery in computational neuroscience: that simple-cell response properties emerge from efficient coding principles \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent}. Much of the subsequent work in dictionary learning has come out of the signal processing and machine learning communities, whose interest in sparse representations as a practical tool motivates the data-driven approach to inferring optimal dictionaries (see \cite{Zhang15} for a comprehensive review). A popular formulation of the idea is to find a set of $m$ `elementary' signals, at most $k$ of which need be linearly combined to represent each $n$-dimensional signal in a dataset of size $N$, typically where $k < m \ll N$.

Blind source identification first started \cite{sato1975method}.

@@

The main finding of this work is that for any $k < m$, every dictionary satisfying the spark condition \eqref{SparkCondition} from compressed sensing is identifiable from as few as $N = m + m(k-1){m \choose k}$ $k$-sparse linear combinations of its elements up to an error linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $(n, m, k)$ satisfy the nearly-optimal compressed sensing inequality \eqref{CScondition}, in almost all cases the dictionary learning problem is well-posed given enough data (Cor.~\ref{ProbabilisticCor}). Interestingly, these guarantees hold even without constraining the recovered matrix to satisfy the spark condition. Imposing this constraint does, however, allow us to state a modification of our main result that applies to the case where only an upper bound on the number of dictionary elements is known (Thm.~\ref{DeterministicUniquenessTheorem2}). We hope the explicit, algorithm-independent criteria we provide here may serve as a useful tool in future analyses of dictionary learning procedures.

We state the problem considered here more precisely as follows. Fix a dictionary represented as the columns $\mathbf{A}_j$ ($j = 1,\ldots,m$) of a matrix $\mathbf{A} \in \mathbb R^{n \times m}$ and suppose a dataset $Z$ consists of measurements:
\begin{align}\label{LinearModel}
\mathbf{z}_i = \mathbf{A}\mathbf{x}_i + \mathbf{n}_i,\indent \text{$i=1,\ldots,N$},
\end{align}
for $k$-\emph{sparse} $\mathbf{x}_i \in \mathbb{R}^m$ having at most $k$ nonzero entries and \emph{noise} $\mathbf{n}_i \in \mathbb{R}^n$, with bounded norm $\| \mathbf{n}_i \|_2 \leq  \eta$ representing our combined worst-case uncertainty in  measuring $\mathbf{A}\mathbf{x}_i$.

\begin{problem}[Sparse linear coding]\label{InverseProblem}
Find a matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^m$ such that $\|\mathbf{z}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \eta$ for all $i$.
\end{problem}

@@

Previous work \cite{li2004analysis, Georgiev05, Aharon06, Hillar15} on the noiseless case $\eta = 0$ has shown that the solution (when it exists) is indeed unique in the sense described above provided the $\mathbf{x}_i$ are sufficiently diverse and the matrix $\mathbf{A}$ satisfies the \textit{spark condition}:
\begin{align}\label{SparkCondition}
\mathbf{A}\mathbf{x}_1 = \mathbf{A}\mathbf{x}_2 \implies \mathbf{x}_1 = \mathbf{x}_2,\indent \text{for all $k$-sparse } \mathbf{x}_1, \mathbf{x}_2.
\end{align}
%
which would be, in any case, a necessary condition for uniqueness if we knew only that the $\mathbf{x}_i$ are $k$-sparse. We introduce the following terminology to handle $\eta > 0$. 
\begin{definition}
Fix $Y = \{ \mathbf{y}_1, \ldots, \mathbf{y}_N\} \subset \mathbb{R}^n$. We say $Y$ has a \emph{$k$-sparse representation in $m$ dimensions} if there exists a matrix $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ such that $\mathbf{y}_i = \mathbf{A}\mathbf{x}_i$ for all $i$. %Equivalently, $\mathbf{Y} = \mathbf{A}\mathbf{X}$ for $\mathbf{Y}$ and $\mathbf{X}$ having the $\mathbf{y}_i$ and $\mathbf{x}_i$ as their columns, respectively. 
This representation is \emph{stable} if for every $\delta_1, \delta_2 \geq 0$, there exists $\varepsilon = \varepsilon(\delta_1, \delta_2) \geq 0$ (with $\varepsilon > 0$ when  $\delta_1, \delta_2 > 0$) such that if a matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^m$ satisfy
\begin{align*}
\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon \ \ \text{for all $i$}
\end{align*}
%
then there is some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that for all $i, j$:
\begin{align}\label{def1}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq \delta_1 \ \ \text{and} \ \ \|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 \leq \delta_2.
\end{align}
\end{definition}
% Kato mentions a distance between n-tuples that minimizes the maximum difference over all possible orderings of one of the n-tuples.
% Make just 'sparse representation in m-dimensions? (i.e. with k unspecified?)

@@

In the next section, we give precise statements of our main results, which include an explicit form for $\varepsilon(\delta_1, \delta_2)$. We then prove our main theorem (Thm.~\ref{DeterministicUniquenessTheorem}) in Sec.~\ref{DUT} while stating some additional definitions and lemmas required for the proof, including our main tool from combinatorial matrix analysis (Lem.~\ref{MainLemma}). Our proof is a refinement of the arguments in \cite{Hillar15} to handle noise and to reduce the sufficient number of samples for all $k < m$ from $N=k{m \choose k}^2$ to $N = m + m(k-1){m \choose k}$. All other proofs are relegated to the Supplemental. Finally, several applications are considered in our discussion, Sec.~\ref{Discussion}.

@@

Before stating our main results, we explain how the spark condition \eqref{SparkCondition} relates to the \emph{lower bound} \cite{Grcar10} of $\mathbf{A}$, which is the largest number $\alpha$ such that \mbox{$\|\mathbf{A}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$. By compactness, every injective linear map has a nonzero lower bound; hence, if $\mathbf{A}$ satisfies \eqref{SparkCondition}, then every submatrix formed from $2k$ of its columns or less has a nonzero lower bound. We therefore define the following domain-restricted lower bound of $\mathbf{A}$:
\begin{align*}
L_k(\mathbf{A}) := \frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{A}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
\end{align*} 
Clearly, $L_k(\mathbf{A}) \geq L_{k'}(\mathbf{A})$ whenever $k < k'$, and for any $\mathbf{A}$ satisfying \eqref{SparkCondition}, we have $L_{k'}(\mathbf{A}) > 0$ for all $k' \leq 2k$. Note that subtracting $L_k(\mathbf{A})$ from one yields what is known in the compressed sensing literature as the (asymmetric) lower restricted isometry constant \cite{Blanchard2011, Foucart2009}.

We let $[m]$ denote the set $\{1, \ldots, m\}$ and $2^{[m]}$ ts power set, and we write ${[m] \choose k}$ to denote the set of subsets of $[m]$ of size $k$. Let $\text{Span}\{\mathbf{v}_1, \ldots, \mathbf{v}_\ell\}$ be the $\mathbb{R}$-linear span of vectors $\mathbf{v}_1, \ldots, \mathbf{v}_\ell$. A vector $\mathbf{x}$ is then said to be \emph{supported} on $S \subseteq [m]$ when $\mathbf{x} \in \text{Span} \{\mathbf{e}_j\}_{j\in S}$, where $\mathbf{e}_j$ are the standard $m$-dimensional basis vectors We denote by $\mathbf{M}_j$ the $j$th column of a matrix $\mathbf{M}$ and similarly by $\mathbf{M}_S$ the sub-matrix formed by the columns of $\mathbf{M}$ indexed by $S$. We interpret $\text{Span}\{\mathbf{M}_\emptyset\} = \{\textbf{0}\}$, where $\emptyset$ is the empty set. We say a set system $E \subseteq 2^{[m]}$ is of rank $k$ when $|S| \leq k$ for every $S \in E$ (and $k$-uniform when equality holds). We say $E$ is \emph{regular} when every element of $[m]$ is contained in exactly $\ell$ elements of $E$ for some $\ell > 0$ (for given $\ell$, we say $E$ is $\ell$-regular).

@@

[State this without reference to special set $E$]
\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix integers $n$ and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} and that \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} include at least \mbox{$(k-1){m \choose k}+1$} vectors in general linear position supported on each set in some regular $E \subseteq {[m] \choose k}$ satisfying the singleton intersection property. Then there exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$. %\frac{L_2(\mathbf{A})}{\sqrt{2}}C_1^{-1}$:

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ for which for some $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m}$ we have \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
\begin{align}\label{Cstable}
\|\mathbf{A}_j-(\mathbf{B}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon \ \ \text{for all $j \in [m]$}
\end{align}
%
for some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$.\footnote{We note that the condition $\varepsilon < L_2(\mathbf{A})$ above is necessary. When \mbox{$\mathbf{A}$ = $I$} and $\mathbf{x}_i = \mathbf{e}_i$, it turns out that $C_1 = 1$ and there is a $\mathbf{B}$ and $1$-sparse $\mathbf{\hat x}_i$ with $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$ violating \eqref{Cstable}.}
Furthermore if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$ then $\mathbf{B}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$ and we in fact have:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,
 \ \ \text{for all $i \in [N]$}.
\end{align}
\end{theorem}

For clarity of exposition, we delay explicitly defining the (effective) constant $C_1 = C_1(\mathbf{A}, \{\mathbf{x}_i\}_{i=1}^N, E)$ until Section \ref{DUT} (\eqref{Cdef}), since it requires additional definitions not critical to the statement of our results.

@@

%\begin{proposition}
%If $E$ is an $\ell$-regular set system of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$. 
%\end{proposition}
Note that if $E$ is an $\ell$-regular set system of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$.
This is easy to see by noting that $|E|k \geq \sum_{S \in E}|S| = \sum_{i \in [m]} \deg(i) = \ell m$. In certain cases there exist simple constructions achieving this minimum; for example, when $k = \sqrt{m}$ one can take $E$ to be the rows and columns formed by arranging the elements of $[m]$ in a square grid.

@@

Since $f$ is a real analytic function, it is enough to show that at least \emph{one} substitution of real numbers satisfies $f(\mathbf{A}) \neq 0$ to conclude that its zeroes form a set with measure zero. Hence, \emph{every} $n \times m$ matrix $\mathbf{A}$ outside such a set satisfies \eqref{SparkCondition} provided \eqref{CScondition} holds for a value of $\gamma$ for \emph{some} distribution. We remark that the precise relationship between $m$, $n$, and $k$ guaranteeing that $f$ is not identically zero is still an open problem in real algebraic geometry [**TODO** cite recent work?]. In any case, we set $\gamma_0$ to be the smallest known such $\gamma$.

@@

\begin{theorem}\label{DeterministicUniquenessTheorem2}
Fix integers $n$ and $k < m \leq m'$ and matrices $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ and $\mathbf{B} \in \mathbb{R}^{n \times m'}$ both satisfying \eqref{SparkCondition}. Suppose \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} include at least \mbox{$(k-1){m' \choose k}+1$} $k$-sparse vectors in general linear position supported on each set in some $k$-uniform $E \subseteq 2^{[m]}$ satisfying the singleton intersection property. Then there exists a constant $C_3 > 0$ for which the following holds.

If for some $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m'}$ and $\varepsilon < L_2(\mathbf{A}) / C_3$ we have \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$, then:
\begin{align}\label{Cstablem'}
\|\mathbf{A}_j-(\mathbf{B}_J\mathbf{PD})_j\|_2 \leq C_3\varepsilon \ \ \text{for all $j \in [m]$}
\end{align}
%
for some $J \in {[m'] \choose m}$, permutation matrix $\mathbf{P}$, and invertible diagonal matrix $\mathbf{D}$.
\end{theorem}

In other words, the columns of $B$ contain (up to noise, after appropriate scaling) the columns of the original dictionary $\mathbf{A}$. Similarly, one can show by the same arguments at the beginning of Sec.~\ref{DUT} that the $\mathbf{\hat x}_i$ contain the original codes $\mathbf{x}_i$. The constant $C_3$ here is expression (\ref{Cdef2}) from the proof of Thm.~\ref{DeterministicUniquenessTheorem2}. Note that, in contrast to Thm.~\ref{DeterministicUniquenessTheorem}, this constant is dependent on $\mathbf{B}$; hence, \eqref{Cstablem'} holds for \emph{all} matrices $\mathbf{B} \in \mathbb{R}^{n \times m'}$ satisfying the spark condition only in the case $\varepsilon = 0$.

>>>>

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{C.G. proved Theorem 1. C.H. proved Theorem 2.}
\authordeclaration{The authors declare no conflict of interest.}
%\equalauthors{\textsuperscript{1}C.G. proved Theorem 1 and C.H. proved Theorem 2.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: cjg@berkeley.edu, cav@mac.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Dictionary learning methods for sparse coding have exposed underlying structure in many kinds of natural signals.
However, universal theorems guaranteeing the statistical consistency of inference in this model are lacking.
Here, we prove that for almost all diverse enough datasets generated from the sparse coding model, latent dictionaries and sparse codes are uniquely determined up to an error commensurate with measurement noise.
Applications are given to data analysis, neuroscience, and engineering.
\end{abstract}

@@

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{D}ictionary learning methods for sparse coding have been under active development ever since their application to natural images brought about an exciting discovery in computational neuroscience: that simple-cell response properties emerge from efficient coding principles \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent}. Much of the subsequent work in dictionary learning has come out of the signal processing and machine learning communities, which have found sparse representations to be a powerful tool for the manipulation of signals lacking much known analytic structure (see \cite{Zhang15} for a comprehensive review). This motivates the data-driven approach to inferring optimal dictionaries for a wide variety of signal classes for which, traditionally, representations have been constructed using methods of harmonic analysis (e.g. the Fourier basis, wavelets, curvelets, etc.). A popular formulation of the idea is to find a set of $m$ `elementary' waveforms, at most $k$ of which need be linearly combined to represent each $n$-dimensional signal in a dataset of size $N$, typically where $k < m \ll N$. 
%\dropcap{D}ictionary learning methods for sparse coding have been under active development ever since their application to natural images brought about an exciting discovery in computational neuroscience: that simple-cell response properties emerge from efficient coding principles \cite{Olshausen96, hurri1996image, bell1997independent, van1998independent}. Much of the subsequent work in dictionary learning has come out of the signal processing and machine learning communities, whose interest in sparse representations as a practical tool motivates the data-driven approach to inferring optimal dictionaries (see \cite{Zhang15} for a comprehensive review). A popular formulation of the idea is to find a set of $m$ `elementary' signals, at most $k$ of which need be linearly combined to represent each $n$-dimensional signal in a dataset of size $N$, typically where $k < m \ll N$.

@@

The main finding of this work is that for any $k < m$, every dictionary satisfying the spark condition \eqref{SparkCondition} from compressive sensing is identifiable from as few as $N = m + m(k-1){m \choose k}$ sparse linear combinations of its elements up to an error linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $(n, m, k)$ satisfy the nearly-optimal compressive sensing inequality \eqref{CScondition}, in almost all cases the dictionary learning problem is well-posed given enough data (Cor.~\ref{ProbabilisticCor}). Interestingly, these guarantees hold even without constraining the recovered matrix to satisfy the spark condition. Imposing this constraint does, however, allow us to state a modification of our main result that applies to the case where only an upper bound on the number of dictionary elements is known (Thm.~\ref{DeterministicUniquenessTheorem2}). We hope the explicit, algorithm-independent criteria we provide here may serve as a useful tool in future analyses of dictionary learning procedures.

We state the problem considered here more precisely as follows. Fix a dictionary represented as the columns $\mathbf{A}_j$ of a matrix $\mathbf{A} \in \mathbb R^{n \times m}$ and suppose a dataset $Z$ consists of measurements:
\begin{align}\label{LinearModel}
\mathbf{z}_i = \mathbf{A}\mathbf{x}_i + \mathbf{n}_i,\indent \text{$i=1,\ldots,N$},
\end{align}
for $k$-\emph{sparse} $\mathbf{x}_i \in \mathbb{R}^m$ having at most $k$ nonzero entries and \emph{noise} $\mathbf{n}_i \in \mathbb{R}^n$, with bounded norm $\| \mathbf{n}_i \|_2 \leq  \eta$ representing our combined worst-case uncertainty in  measuring $\mathbf{A}\mathbf{x}_i$.

\begin{problem}[Sparse Linear Coding]\label{InverseProblem}
Find a matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^m$ such that $\|\mathbf{z}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \eta$ for all $i$.
\end{problem}

@@

Previous work \cite{li2004analysis, Georgiev05, Aharon06, Hillar15} on the noiseless case $\eta = 0$ has shown that the solution (when it exists) is indeed unique in the sense described above provided the $\mathbf{x}_i$ are sufficiently diverse and the matrix $\mathbf{A}$ satisfies the \textit{spark condition}:
\begin{align}\label{SparkCondition}
\mathbf{A}\mathbf{x}_1 = \mathbf{A}\mathbf{x}_2 \implies \mathbf{x}_1 = \mathbf{x}_2,\indent \text{for all $k$-sparse } \mathbf{x}_1, \mathbf{x}_2.
\end{align}
%
which would be a necessary condition for uniqueness if we knew only that the $\mathbf{x}_i$ are $k$-sparse. We introduce the following terminology to handle $\eta > 0$. 
\begin{definition}
Fix $Y = \{ \mathbf{y}_1, \ldots, \mathbf{y}_N\} \subset \mathbb{R}^n$. We say $Y$ has a \emph{$k$-sparse representation in $m$ dimensions} if there exists a matrix $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ such that $\mathbf{y}_i = \mathbf{A}\mathbf{x}_i$ for all $i$. %Equivalently, $\mathbf{Y} = \mathbf{A}\mathbf{X}$ for $\mathbf{Y}$ and $\mathbf{X}$ having the $\mathbf{y}_i$ and $\mathbf{x}_i$ as their columns, respectively. 
This representation is \emph{stable} if for every $\delta_1, \delta_2 \geq 0$, there exists $\varepsilon = \varepsilon(\delta_1, \delta_2) \geq 0$ (with $\varepsilon > 0$ when  $\delta_1, \delta_2 > 0$) such that if a matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^m$ satisfy
\begin{align*}
\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon \ \ \text{for all $i$}
\end{align*}
%
then there is some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that for all $i = 1,\ldots,N$ and $j = 1,\ldots,m$:
\begin{align}\label{def1}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq \delta_1 \ \ \text{and} \ \ \|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 \leq \delta_2.
\end{align}
\end{definition}
% Kato mentions a distance between n-tuples that minimizes the maximum difference over all possible orderings of one of the n-tuples.
% Make just 'sparse representation in m-dimensions? (i.e. with k unspecified?)

@@

In the next section, we give precise statements of our main results, which include an explicit form for $\varepsilon(\delta_1, \delta_2)$. We then prove our main theorem (Thm.~\ref{DeterministicUniquenessTheorem}) in Sec.~\ref{DUT} while stating some additional definitions and lemmas required for the proof, including our main tool from combinatorial matrix analysis (Lem.~\ref{MainLemma}). Our proof is a refinement of the arguments in \cite{Hillar15} to handle noise and to reduce the sufficient number of samples for all $k < m$ from $N=k{m \choose k}^2$ to $N = m + m(k-1){m \choose k}$. All other proofs are relegated to the Appendix. Finally, several applications are considered in our discussion, Sec.~\ref{Discussion}.

@@

Before stating our results, we explain how the spark condition \eqref{SparkCondition} relates to the \emph{lower bound} \cite{Grcar10} of $\mathbf{A}$, written $L(\mathbf{A})$, which is the largest number $\alpha$ such that \mbox{$\|\mathbf{A}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$. By compactness, every injective linear map has a nonzero lower bound; hence, if $\mathbf{A}$ satisfies \eqref{SparkCondition}, then every submatrix formed from $2k$ of its columns or less has a nonzero lower bound. We therefore define the following domain-restricted lower bound of $\mathbf{A}$:
\begin{align*}
L_k(\mathbf{A}) := \frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{A}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
\end{align*} 
Clearly, $L_k(\mathbf{A}) \geq L_{k'}(\mathbf{A})$ whenever $k < k'$, and for any $\mathbf{A}$ satisfying \eqref{SparkCondition}, we have $L_{k'}(\mathbf{A}) > 0$ for all $k' \leq 2k$. Note that subtracting $L_k(\mathbf{A})$ from one yields what is known in the compressive sensing literature as the (asymmetric) lower restricted isometry constant \cite{Blanchard2011, Foucart2009}.

We let $[m]$ denote the set $\{1, \ldots, m\}$, let $2^{[m]}$ denote its power set, and write ${[m] \choose k}$ to denote the set of subsets of $[m]$ of size $k$. Let $\text{Span}\{\mathbf{v}_1, \ldots, \mathbf{v}_\ell\}$ be the $\mathbb{R}$-linear span of vectors $\mathbf{v}_1, \ldots, \mathbf{v}_\ell$. A vector $\mathbf{x}$ is then said to be \emph{supported} on $S \subseteq [m]$ when $\mathbf{x} \in \text{Span} \{\mathbf{e}_j\}_{j\in S}$, where $\mathbf{e}_j$ are the standard $m$-dimensional basis vectors We denote by $\mathbf{M}_j$ the $j$th column of a matrix $\mathbf{M}$ and similarly by $\mathbf{M}_S$ the sub-matrix formed by the columns of $\mathbf{M}$ indexed by $S$. We interpret $\text{Span}\{\mathbf{M}_\emptyset\} = \{\textbf{0}\}$, where $\emptyset$ is the empty set. We say a set system $E \subseteq 2^{[m]}$ is of rank $k$ when $|S| \leq k$ for every $S \in E$ (and $k$-uniform when equality holds). We say $E$ is \emph{regular} when every element of $[m]$ is contained in exactly $\ell$ elements of $E$ for some $\ell > 0$ (for given $\ell$, we say $E$ is $\ell$-regular).

@@

\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix integers $n$ and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} and that \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} include at least \mbox{$(k-1){m \choose k}+1$} vectors in general linear position supported on each set in some regular $E \subseteq {[m] \choose k}$ satisfying the singleton intersection property. Then there exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$. %\frac{L_2(\mathbf{A})}{\sqrt{2}}C_1^{-1}$:

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ for which for some $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m}$ we have \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
\begin{align}\label{Cstable}
\|\mathbf{A}_j-(\mathbf{B}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon \ \ \text{for all $j \in [m]$}
\end{align}
%
for some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$.\footnote{We note that the condition $\varepsilon < L_2(\mathbf{A})$ above is necessary. When \mbox{$\mathbf{A}$ = $I$} and $\mathbf{x}_i = \mathbf{e}_i$ it turns out that $C_1 = 1$ and there is a $\mathbf{B}$ and $1$-sparse $\mathbf{\hat x}_i$ with $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$ violating \eqref{Cstable}.}
Furthermore if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$ then $\mathbf{B}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$ and we in fact have:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq \frac{\varepsilon}{\sqrt{2k}} \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right).
 \ \ \text{for all $i \in [N]$}.
\end{align}
\end{theorem}

For clarity of exposition, we delay explicitly defining the constant $C_1 = C_1(\mathbf{A}, \{\mathbf{x}_i\}_{i=1}^N, E)$ until Section \ref{DUT} (\eqref{Cdef}), since it requires additional definitions not critical to the statement of our results.

@@

\begin{proposition}
If $E$ is an $\ell$-regular set system of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$. 
\end{proposition}
This is easy to see by noting that $|E|k \geq \sum_{S \in E}|S| = \sum_{i \in [m]} \deg(i) = \ell m$. In certain cases there exist simple constructions achieving this minimum; for example, when $k = \sqrt{m}$ one can take $E$ to be the rows and columns formed by arranging the elements of $[m]$ in a square grid.

@@

Since $f$ is a real analytic function, it is enough to show that at least \emph{one} substitution of real numbers satisfies $f(\mathbf{A}) \neq 0$ to conclude that its zeroes form a set with measure zero. Hence, \emph{every} $n \times m$ matrix $\mathbf{A}$ outside a set of measure zero satisfies \eqref{SparkCondition} provided \eqref{CScondition} holds for a value of $\gamma$ for \emph{some} distribution. We remark that the precise relationship between $m$, $n$, and $k$ guaranteeing that $f$ is not identically zero is still an open problem in real algebraic geometry [**TODO** cite recent work?]. In any case, we set $\gamma_0$ to be the smallest known such $\gamma$.

@@

\begin{theorem}\label{DeterministicUniquenessTheorem2}
Fix integers $n$ and $k < m \leq m'$ and matrices $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ and $\mathbf{B} \in \mathbb{R}^{n \times m'}$ both satisfying \eqref{SparkCondition}. Suppose \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} include at least \mbox{$(k-1){m' \choose k}+1$} vectors in general linear position supported on each set in some $k$-uniform $E \subseteq 2^{[m]}$ satisfying the singleton intersection property. Then there exists a constant $C_3 > 0$ for which the following holds.

If for some $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m'}$ and $\varepsilon < L_2(\mathbf{A}) / C_3$ we have \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$, then:
\begin{align}\label{Cstablem'}
\|\mathbf{A}_j-(\mathbf{B}_J\mathbf{PD})_j\|_2 \leq C_3\varepsilon \ \ \text{for all $j \in [m]$}
\end{align}
%
for some $J \in {[m'] \choose m}$, permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$.
\end{theorem}

In other words, the columns of $B$ contain (up to noise, after appropriate scaling) the columns of the original dictionary $\mathbf{A}$. Similarly, one can show by the same arguments at the beginning of Sec.~\ref{DUT} that the $\mathbf{\hat x}_i$ contain the original codes $\mathbf{x}_i$. The constant $C_3$ here is expression (\ref{Cdef2}) from the proof of Thm.~\ref{DeterministicUniquenessTheorem2}. Note that, in contrast to Thm.~\ref{DeterministicUniquenessTheorem}, this constant is is dependent on $\mathbf{B}$; hence \eqref{Cstablem'} holds for \emph{all} matrices $\mathbf{B} \in \mathbb{R}^{n \times m'}$ satisfying the spark condition only in the case $\varepsilon = 0$.
